---
title: "Analysing the ChickWeight Dataset"
subtitle: 'MSc in Statistics 2025/26, Imperial College London'
author: "Jegan G G S"
format:
  html:
    toc: true
    highlight: tango
    self-contained: true
  pdf: default
format-links: false
bibliography: ref2.bib
---

## Objective
Adapt the the simulations and analyses of the palmer penguins dataset to instead use the `ChickWeight` dataset that comes with R. The ChickWeight dataset tracks a number of chicks, their weights measured at regular intervals, and the diets that they were given.

```{r}
#| echo: false
chicks <- ChickWeight 
```

```{r}
#| echo: false
#| output: false
#Number of chicks:
length(unique(chicks$Chick))

#Number of diets
length(unique(chicks$Diet))
```
Number of chicks: 50

Number of different diets: 4


First, we plot the weight of the chicks vs their age: 
```{r}
plot(x = chicks$Time, 
     y = chicks$weight,
      main='Age vs Weight',
      xlab = 'Age',
      ylab = 'Weight')
```

Now, we create a table to observe how many chicks there are for each diet:
```{r}
#CORRECT THIS LATER
table(chicks$Diet)
```

Now, we plot a figure to observe the progress of the weights of the chicks for each diet:
```{r}
plot(
  x = chicks$Time,
  y = chicks$weight, 
  main = "Chick Weight vs Age",
  xlab = "Age",
  ylab = "Weight",
  col = chicks$Diet,
  type='p')
```

We also do boxplots by diet to compare the weights of the chicks following each diet:
```{r}
plot(
  x = chicks$Diet,
  y = chicks$weight, 
  main = "Chick Weight vs Diet",
  xlab = "Diet",
  ylab = "Weight",
  col = c(2,3,4,5))
abline(h = mean(chicks$weight,na.rm = TRUE), col = 2, lwd = 2)
```